---
title: "Bird Joins"
author: "Erica Dale"
date: "2022-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
bird_observations <- read_csv(here("data", "52_pp52_birds_1.csv"))
sites <- read_csv(here("data", "52_pp52_sites_1.csv"))
surveys <- read_csv(here("data", "52_pp52_surveys_1.csv"))
taxalist <- read_csv(here("data", "52_pp52_taxalist_1.csv"))
```

Create a subset of bird_observations called birds_subset that only contains observations for birds with species id “BHCO” and “RWBL”, and from sites with site ID “LI-W” and “NU-C”
```{r}
birds_subset <- bird_observations |> 
  filter(species_id %in% c("BHCO", "RWBL"),
         site_id %in% c("LI-W", "NU-C"))
```


### Left Join Practice
```{r}
birds_left <- left_join(birds_subset, sites, by = "site_id") |> 
  left_join(taxalist, by = "species_id")
```

### Full Join Practice
```{r}
full_join <- full_join(birds_subset, sites, by = "site_id")
```

### Inner Join Practice
```{r}
birds_inner <- inner_join(birds_subset, taxalist, by = "species_id")
```

## Task 2
### Data wrangling and visualization

Starting with your birds object, rename the notes column to bird_obs_notes (so this doesn’t conflict with notes in the surveys dataset:
```{r}
bird_observations <- bird_observations |> 
  rename(bird_obs_notes = notes)
```

Then, create a subset that contains all observations in the birds dataset, joins the taxonomic, site and survey information to it, and is finally limited to only columns survey_date, common_name, park_name, and bird_count. You can decide the order that you want to create this in (e.g. limit the columns first, then join, or the other way around).:
Use lubridate::month() to add a new column called survey_month, containing only the month number. Then, convert the month number to a factor (again within mutate())
Learn a new function on your own! Use dplyr::relocate() to move the new survey_month column to immediately after the survey_date column. You can do this in a separate code chunk, or pipe straight into it from your existing code.



```{r}
all_subset <- bird_observations |> 
  left_join(taxalist, by = "species_id") |> 
  left_join(sites, by = "site_id") |> 
  left_join(surveys, by = "site_id") |> 
  select(survey_date, common_name, park_name, bird_count) |> 
  mutate(survey_month = lubridate::month(survey_date)) |> 
  mutate(survey_month = as.factor(survey_month)) |> 
  relocate(survey_month, .after = survey_date)


class(all_subset$survey_month)
```

Find the total number of birds observed by park and month (i.e., you’ll group_by(park_name, survey_month))
Filter to only include parks “Lindo”, “Orme”, “Palomino” and “Sonrisa”


```{r}
totalbirds <- all_subset |> 
  group_by(park_name, survey_month) |> 
  summarize(bird_sum = sum(bird_count)) |> 
  filter(park_name %in% c("Lindo", "Orme", "Palomino", "Sonrisa"))
```


